---
title: "Mice"
author: "Michel de Lange"
date: "8/7/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, results='HIDE')
```

# Introduction

This is my first attempt to look at the mouse data. I have only looked at a tiny part of the spreadsheet: The CE19pos_B220 data. If I understand it correctly, we have measurements at timepoints 0,3,10,17,24 and 31. For each mouse, we have the number of live cells, and then the number of CD19pos_B220 cells. What matters is the proportion of CD19..220  to live cells.
Mice come in three types: WT, Pound and Plt, and obviously we have male and female mice. We are interested in comparing Plt vs WT, and Pound vs. WT. We do not compare male mice with female mice. So the question is: Is there a difference in said proportion of cells at the different timepoints, in those groups of mice?

# the data

The following table shows the number of mice, by type and gender:

```{r }
mice <- read.csv("/Users/micheldelange/Documents/mice/data/mice.csv",header=TRUE)[1:40,]
table(mice$Animal_type)
```

So we have a total of `r sum(table(mice$Animal_type))` mice, 20 of each gender. Let's look at the proportions of said cells, at the 6 timepoints: First the girl mice:


```{r }


#tail(mice$Data.Set)
#table(mice$Animal_type)

mice$CD19pos_B220pos_0_prop <-  mice$CD19pos_B220pos_0 / mice$Live_cells_0
mice$CD19pos_B220pos_3_prop <-  mice$CD19pos_B220pos_3 / mice$Live_cells_3
mice$CD19pos_B220pos_10_prop <-  mice$CD19pos_B220pos_10 / mice$Live_cells_10
mice$CD19pos_B220pos_17_prop <-  mice$CD19pos_B220pos_17/ mice$Live_cells_17
mice$CD19pos_B220pos_24_prop <-  mice$CD19pos_B220pos_24 / mice$Live_cells_24
mice$CD19pos_B220pos_31_prop <-  mice$CD19pos_B220pos_31 / mice$Live_cells_31



par(mfrow=c(1,1))
#boxplot(mice$Animal_type,mice$CD19pos_B220pos_0_perc)

female <- mice[grepl('female',mice$Animal_type),]
male   <- mice[grepl('_male',mice$Animal_type),] 

female$Animal_type = factor(female$Animal_type)
male$Animal_type = factor(male$Animal_type)
female$plt2 <- grepl('PLT2',female$Animal_type)
female$pound <- grepl('Pound_',female$Animal_type)
male$plt2 <- grepl('PLT2',male$Animal_type)
male$pound <- grepl('Pound_',male$Animal_type)


#table(female$Animal_type)
#table(male$Animal_type)


par(mfrow=c(1,2))


  
## same thing, by animal type
male_wt <- mice[grepl('WT_male',mice$Animal_type),] 
male_pound <- mice[grepl('Pound_male',mice$Animal_type),] 
male_plt2 <- mice[grepl('PLT2_male',mice$Animal_type),] 

female_wt <- mice[grepl('WT_female',mice$Animal_type),] 
female_pound <- mice[grepl('Pound_female',mice$Animal_type),] 
female_plt2 <- mice[grepl('PLT2_female',mice$Animal_type),] 


par(mfrow=c(1,3))


## female mice


{
plot(' ',xlim=c(0,31),ylim=c(0,0.5),main='WT_female',ylab='CD19pos_B220',xlab='time')
for (mouse in seq(1:6)) {
  
  lines(c(0,3,10,17,24,31),
        c(female_wt$CD19pos_B220pos_0_prop[mouse],
          female_wt$CD19pos_B220pos_3_prop[mouse],
          female_wt$CD19pos_B220pos_10_prop[mouse],
          female_wt$CD19pos_B220pos_17_prop[mouse],
          female_wt$CD19pos_B220pos_24_prop[mouse],
          female_wt$CD19pos_B220pos_31_prop[mouse]
        ),
        pch='.',col=rainbow(6)[mouse])
}
}

{
plot(' ',xlim=c(0,31),ylim=c(0,0.5),main='Pound_female',ylab='CD19pos_B220',xlab='time')
for (mouse in seq(1:9)) {
  lines(c(0,3,10,17,24,31),
        c(female_pound$CD19pos_B220pos_0_prop[mouse],
          female_pound$CD19pos_B220pos_3_prop[mouse],
          female_pound$CD19pos_B220pos_10_prop[mouse],
          female_pound$CD19pos_B220pos_17_prop[mouse],
          female_pound$CD19pos_B220pos_24_prop[mouse],
          female_pound$CD19pos_B220pos_31_prop[mouse]
        ),
        pch='.',col=rainbow(9)[mouse])
}
}

{
plot(' ',xlim=c(0,31),ylim=c(0,0.5),main='plt2_female',ylab='CD19pos_B220',xlab='time')
for (mouse in seq(1:5)) {
  lines(c(0,3,10,17,24,31),
        c(female_plt2$CD19pos_B220pos_0_prop[mouse],
          female_plt2$CD19pos_B220pos_3_prop[mouse],
          female_plt2$CD19pos_B220pos_10_prop[mouse],
          female_plt2$CD19pos_B220pos_17_prop[mouse],
          female_plt2$CD19pos_B220pos_24_prop[mouse],
          female_plt2$CD19pos_B220pos_31_prop[mouse]
        ),
        pch='.',col=rainbow(5)[mouse])
}
}

```

And now the boy mice:

```{r }
par(mfrow=c(1,3))
{
plot(' ',xlim=c(0,31),ylim=c(0,0.5),main='WT_male',ylab='CD19pos_B220',xlab='time')
for (mouse in seq(1:4)) {
  
  lines(c(0,3,10,17,24,31),
        c(male_wt$CD19pos_B220pos_0_prop[mouse],
          male_wt$CD19pos_B220pos_3_prop[mouse],
          male_wt$CD19pos_B220pos_10_prop[mouse],
          male_wt$CD19pos_B220pos_17_prop[mouse],
          male_wt$CD19pos_B220pos_24_prop[mouse],
          male_wt$CD19pos_B220pos_31_prop[mouse]
        ),
        pch='.',col=rainbow(4)[mouse])
}
}

{
plot(' ',xlim=c(0,31),ylim=c(0,0.5),main='Pound_male',ylab='CD19pos_B220',xlab='time')
for (mouse in seq(1:7)) {
  lines(c(0,3,10,17,24,31),
        c(male_pound$CD19pos_B220pos_0_prop[mouse],
          male_pound$CD19pos_B220pos_3_prop[mouse],
          male_pound$CD19pos_B220pos_10_prop[mouse],
          male_pound$CD19pos_B220pos_17_prop[mouse],
          male_pound$CD19pos_B220pos_24_prop[mouse],
          male_pound$CD19pos_B220pos_31_prop[mouse]
        ),
        pch='.',col=rainbow(7)[mouse])
}
}

{
plot(' ',xlim=c(0,31),ylim=c(0,0.5),main='Plt2_male',ylab='CD19pos_B220',xlab='time')
for (mouse in seq(1:9)) {
  lines(c(0,3,10,17,24,31),
        c(male_plt2$CD19pos_B220pos_0_prop[mouse],
          male_plt2$CD19pos_B220pos_3_prop[mouse],
          male_plt2$CD19pos_B220pos_10_prop[mouse],
          male_plt2$CD19pos_B220pos_17_prop[mouse],
          male_plt2$CD19pos_B220pos_24_prop[mouse],
          male_plt2$CD19pos_B220pos_31_prop[mouse]
        ),
        pch='.',col=rainbow(9)[mouse])
}
}
```


There seems to be a lot of individuality in mice. Whatever relationship there is, it is certainly not linear. Let's look at boxplots. Here we see the information grouped together by type and gender, ignoring how it goes for an individual mouse. First, the girl mice again:

```{r }


plot(female$Animal_type,female$CD19pos_B220pos_0_prop,main='female, time 0')
plot(female$Animal_type,female$CD19pos_B220pos_3_prop,main='female, time 3')
plot(female$Animal_type,female$CD19pos_B220pos_10_prop,main='female, time 10')
plot(female$Animal_type,female$CD19pos_B220pos_17_prop,main='female, time 17')
plot(female$Animal_type,female$CD19pos_B220pos_24_prop,main='female, time 24')
plot(female$Animal_type,female$CD19pos_B220pos_31_prop,main='female, time 31')

```

And now the boy mice:

```{r }

{
plot(male$Animal_type,male$CD19pos_B220pos_0_prop,main='male, time 0')
plot(male$Animal_type,male$CD19pos_B220pos_3_prop,main='male, time 3')
plot(male$Animal_type,male$CD19pos_B220pos_10_prop,main='male, time 10')
plot(male$Animal_type,male$CD19pos_B220pos_17_prop,main='male, time 17')
plot(male$Animal_type,male$CD19pos_B220pos_24_prop,main='male, time 24')
# no data
#plot(male$Animal_type,male$CD19pos_B220pos_31_prop,main='male, time 31')
}

```

